Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
polishing example 3

\begin{subfigure}[t]{.2\textwidth}
  {\small
\begin{verbatim}
int main() {
  int x;
  x = f();
  return g();
}

@@

20 g:    add $1,%sp      ; allocate frame
21       mov -1(%sp),%ra ; violation!
22       sub $1,%sp      ; deallocate frame
23       ret
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{62}
  \MemoryLabel{-10em}{0.75em}{105}
  \\
  {\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \\
  {\tt 4} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{2}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  {\tt 22} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{\bf 5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{36}
  \MemoryLabel{-10em}{0.75em}{\bf 8}
  \MemoryLabel{-6em}{0.75em}{90}
  \\
\end{tabular}
\end{center}

@@

\caption{Another confidentiality violation}
\label{fig:conf2}
\end{figure}

It may be unclear in figure \ref{fig:conf1} why we vary the initial values of {\tt y}
and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
temporal stack safety.
In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
that corresponds to {\tt main}'s variable, {\tt x}. Then {\tt g} is able to access
{\tt y} and hence {\tt x} without even directly reading the stack pointer.

@@

10 f:     add $1,%sp     ; alloc frame
11        cmp stash, $0  ; stash = 0?
12        breq #15       ; branch if so
13        mov -2(%sp),stash ; set stash
14        jmp #17
15        mov stash,-2(%sp) ; get stash
16        mov $0,stash
17        sub $1,%sp     ; dealloc frame
18        ret
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.25\textwidth}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \hspace*{-1.5em}
      \memory[1.2em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-2.2em}{0.75em}{\#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
    \end{tabular}
  \end{subfigure}
  \caption{A control-flow attack}
  \label{fig:controlflow}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{subfigure}{.2\textwidth}
{\small
\begin{verbatim}
int *stash = 0;

int main() {
  int x, y;
  f(x);
  x = 0;
  f(y);
  return x;
}

void f(int &a) {
  if(!stash)
    stash = &a;
  *stash = 5;
  return;
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.49\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $2,%sp      ; allocate frame
2         mov %sp,%ra
3         add -2,%ra      ; ra = &x
4         call f          ; SHARE: %sp-2
5         mov $0,-2(%sp)  ; x = 0
6         mov %sp,%ra
7         add -1,%ra      ; ra = &y
8         call f          ; SHARE: %sp-1
9         mov -2(%sp),%ra ; set return value
10        sub $2,%sp      ; deallocate frame
11        ret

20 f:     cmp stash, $0   ; compare
21        breq #13        ; branch if equal
22        mov %ra,stash   ; stash = &a
23        jmp #15
24        mov $5, (stash) ; *stash = 5
25        ret
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.29\textwidth}
    \begin{tabular}{l l}
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 8} &
      \memoryaddrs[(0)]{12em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}
\caption{A violation with pass-by-reference}
\label{fig:passing}
\end{figure}

>>>>

\begin{figure}
\begin{subfigure}[t]{.2\textwidth}
  {\small
\begin{verbatim}
int main() {
  int x;
  x = f();
  return g();
}

@@

20 g:    add $1,%sp      ; allocate frame
21       mov -1(%sp),%ra ; violation!
22       sub $1,%sp      ; deallocate frame
23       ret
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{3}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-15em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(5\)}
  \\
  {\tt 4} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{1}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-14em}{0.75em}{5}
%    \MemoryLabel{-10em}{0.75em}{\(v_1\)}
    \MemoryLabel{-6em}{0.75em}{5}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  {\tt 22} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-15em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\bf 5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-15em}{0.75em}{\(v_3\)}
  \MemoryLabel{-6em}{0.75em}{\(v_4\)}
%  \MemoryLabel{-2em}{0.75em}{90}
  \\
\end{tabular}
\end{center}

@@

\end{subfigure}
\caption{Another confidentiality violation}
\label{fig:conf2}
\end{figure}

It may be unclear in figure \ref{fig:conf1} why we vary the initial values of {\tt y}
and {\tt z} as well as that of {\tt x}. This is to detect violations of so-called
temporal stack safety.
In Figure~\ref{fig:conf2}, we watch as {\tt f} leaves behind a value in {\tt y}
that later becomes the value of {\tt main}'s variable {\tt x}. Then {\tt g} is able to access
{\tt y} and hence {\tt x} without even directly reading the stack pointer.
In the primary trace, {\tt z} will have the value 5, but in the variant it can have
an arbitrary value $v_4$; since the variant can behave different from the primary,
the policy again signals a confidentiality violation.

@@

10 f:     add $1,%sp     ; alloc frame
11        cmp stash, $0  ; stash = 0?
12        breq #15       ; branch if so
13        mov -2(%sp),stash ; set stash
14        jmp #17
15        mov stash,-2(%sp) ; get stash
16        mov $0,stash
17        sub $1,%sp     ; dealloc frame
18        ret
\end{verbatim}
}
  \end{subfigure}
%
  \begin{subfigure}[t]{.2\textwidth}
    \begin{center}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \hspace*{-1.5em}
      \memory[1.2em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-2.2em}{0.75em}{\#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
    \end{tabular}
    \end{center}
    \vspace{\abovedisplayskip}
  \end{subfigure}
  \caption{A control-flow attack.}
  \label{fig:controlflow}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{subfigure}[t]{.2\textwidth}
{\small
\begin{verbatim}
int *stash = 0;
int main() {
  int x, y;
  f(x);
  x = 0;
  f(y);
  return x;
}

void f(int &a) {
  if(!stash)
    stash = &a;
  *stash = 5;
  return;
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.49\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $2,%sp      ; allocate frame
2         mov %sp,%ra
3         add -2,%ra      ; ra = &x
4         call f          ; SHARE: %sp-2
5         mov $0,-2(%sp)  ; x = 0
6         mov %sp,%ra
7         add -1,%ra      ; ra = &y
8         call f          ; SHARE: %sp-1
9         mov -2(%sp),%ra ; set return value
10        sub $2,%sp      ; deallocate frame
11        ret

20 f:     cmp stash, $0   ; stash = 0? 
21        breq #13        ; branch if so
22        mov %ra,stash   ; stash = &a
23        jmp #15
24        mov $5, (stash) ; *stash = 5
25        ret
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.29\textwidth}
    \begin{center}
    \begin{tabular}{l l}
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 8} &
      \memoryaddrs[(0)]{12em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{center}
\vspace{\abovedisplayskip}
\end{subfigure}
\caption{A violation with pass-by-reference}
\label{fig:passing}
\end{figure}
